---
title: "PacifichakeMSE"
author: "Nis S Jacobsen"
date: "April 30, 2020"
email: "nsja@aqua.dtu.dk"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
# Introduction

This is a description of the 'PacifichakeMSE' package to run management strategy evaluation of Pacific hake. The package sets up a set of parameters, and runs an operating model in R, and fits an estimation in TMB.  
The package runs on a range of dependencies, most notably TMB, which require a developer version of R. For more information on TMB see https://github.com/kaskr/adcomp

To install the 'PacifichakeMSE' package run 
```{r, eval = FALSE}
devtools::install_github('https://github.com/nissandjac/PacifichakeMSE')
```

Load the dependencies 
```{r message=FALSE}
library(TMB)
library(r4ss)
library(devtools)
library(PacifichakeMSE)
library(dplyr)
library(ggplot2)
```

# Operating model 

First we can initialize the parameters of the operating model using the function 'load_data_seasons'
```{r , echo = TRUE}

df <- load_data_seasons(nseason = 4, # Specifies the number of seasons
                        nspace = 2, # Specifies the number of areas 
                        logSDR = 1.4) # Recruitment deviation 

# See 
help("load_data_seasons")

```


