---
title: "Conditioning of Pacific hake operating model"
author: "Nis S Jacobsen (nissandjac@gmail.com)"
date: "November 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

## Introduction 

This file runs the conditioning of the Pacific hake management strategy evaluation operating model. We run a simple 


## Data initialization

Load required packages
```{r message=FALSE, warning=FALSE}
library(TMB)
library(dplyr)
library(reshape2)
library(ggplot2)
library(r4ss)
```
Source the functions required to run the MSE (put into package later)
```{r}
source('load_files_OM.R')
```
Load the assessment data
```{r}
assessment <- read.csv('data/assessment_MLE.csv')
assessment <- assessment[assessment$year > 1965,]
mod <- SS_output(paste(getwd(),'/data/', sep =''), 
                 printstats=FALSE, 
                 verbose = FALSE, 
                 hidewarn = TRUE, 
                 warn = FALSE
                 )

```

Initialize the model parameters. Make a version with no movement or seasons (simple)
```{r}

df <- load_data_seasons(nseason = 4, nspace = 2, bfuture = 0.5, movemaxinit = 0.5, movefiftyinit =8)
df.simple <- load_data_seasons(nseason = 1, nspace = 1, bfuture = 0.5) # Prepare data for operating model
```


Run the model using 'run.agebased.true.catch()'

```{r}
sim.data <- run.agebased.true.catch(df)
sim.data.simple <- run.agebased.true.catch(df.simple)

```

Plot the spawning biomass of the operating model. The diffference in spawning biomass is due to spatial structure 
```{r echo=FALSE}
# Plot the biomass in ggplot 
df.plot <- data.frame(years = c(df$years,df$years,assessment$year), 
                      SSB = c(rowSums(sim.data$SSB.weight),sim.data.simple$SSB.weight,assessment$SSB), 
                      source = c(rep('SSB OM', length(df$years)), 
                                 rep('SSB simple', length(df.simple$years)),
                                 rep('SSB assessment', length(assessment$year))
                                 ))

p1 <- ggplot(data = df.plot, aes(x = years, y = SSB, color = source))+geom_line(size = 2)+theme_classic()
p1

```

Plot the survey data and estimates 

```{r echo=FALSE}
survey.ss <- data.frame(years = mod$cpue$Yr,
                        survey =mod$cpue$Exp,
                        source = 'SS',
                        survsd = NA,
                        kriegsd = NA)

df.plot <- data.frame(years = rep(df$years[df$survey > 1],3),
                      survey = c(df$survey[df$survey > 1],sim.data$survey[sim.data$survey > 1],sim.data.simple$survey[sim.data$survey > 1]),
                      source = rep(c('Survey data','OM output','simple model'), each = length(df$years[df$survey > 1])),
                      survsd= c(df$survey_err[df$flag_survey ==1], 
                                rep(NA,length(df$years[df$survey > 1])), 
                                rep(NA,length(df$years[df$survey > 1]))),
                      kriegsd = c(rep(exp(df$parms$logSDsurv),length(df$years[df$survey > 1])), 
                                  rep(NA,length(df$years[df$survey > 1])), 
                                  rep(NA,length(df$years[df$survey > 1])))
)

df.plot <- rbind(df.plot,survey.ss)

df.plot$survsd <- sqrt(df.plot$survey^2*exp(df.plot$survsd+df.plot$kriegsd-1))

p2 <- ggplot(data = df.plot, aes(x = years, y = survey/1e6, color = source))+
  geom_point(data = df.plot[df.plot$source == 'Survey data',],size = 3)+
  #geom_line(data = df.plot[df.plot$source == 'OM output',], size =2)+
  #geom_line(data = df.plot[df.plot$source == 'SS',], size = 2)+
  geom_line(data = df.plot[df.plot$source != 'Survey data',], size = 1.5)+
  theme_classic()+
  geom_errorbar(aes(ymin=(survey-survsd)/1e6, ymax=(survey+survsd)/1e6))+
  scale_y_continuous(limit = c(0,5), name = 'survey biomass (million t)')+
  scale_x_continuous(name = 'year')

p2
```

Calculate average ages using the calcMeanAge() function
```{r}
source('calcMeanAge.R')

age.comps <- sim.data$age_comps_catch_space
age.comps <- apply(age.comps,c(1,2),sum)/2

am <- calcMeanAge(age.comps,df$age_maxage)
```
Plot the average age in the catch 
```{r echo=FALSE}
age.comps.can <- sim.data$age_comps_catch_space[,,1]
am.can <-calcMeanAge(age.comps.can, df$age_maxage)

age.comps.US <- sim.data$age_comps_catch_space[,,2]
am.US <- calcMeanAge(age.comps.US, df$age_maxage)

# Observed age comps
cps <- read.csv('data/catch_per_sector.csv')

## Calculate d

cps.s <- melt(cps, id.vars = c('year','nfish','nhauls','Country','Catch'),
              measure.vars = rep(paste('X',1:15, sep =''))) # Omit fleet from this df

cps.s <- cps.s %>% 
  group_by(year,Country, variable) %>% 
  summarise(agecomp =weighted.mean(value,Catch))

# Make ages numbers rather than factors
ages <- as.numeric(unlist(strsplit(as.character(cps.s$variable), split = "X")))
ages <- ages[is.na(ages) == 0]

cps.s$age <- ages

cps.am <- cps.s %>% 
  group_by(year,Country) %>% 
  summarise(am = sum((agecomp/100)*age))

cps.am.all <- matrix(NA, df$nyear)

for(i in 1:df$nyear){
  cps.am.all[i] <- sum(df$age[2:16]*df$age_catch[,i])
}

df.am.all <- data.frame(year = df$years, Country = 'All',am = cps.am.all)


df.am <- data.frame(year = rep(df$years,3), 
                    am = c(am,
                           am.can,
                           am.US), 
                    Country = rep(c('All','Can','US'), each = length(df$years)))

df.tmp <- data.frame(year = rep(unique(cps.am$year),2),
                     am = c(cps.am$am[cps.am$Country == 'Can'],
                            cps.am$am[cps.am$Country == 'US']),
                            Country = rep(c('Can obs','US obs'), each = length(unique(cps.am$year)))
                     )

df.am <-rbind(df.am,df.tmp)

p3 <- ggplot(df.am, aes(x = year, y = am, color = Country))+geom_line(size = 1)+
  theme_classic()+
  scale_y_continuous(name = 'Average age in catch', limit = c(2,10))+scale_x_continuous()

p3

```

```{r echo = FALSE}
# png('Age_comps_catch.png', width = 16, height = 12, res = 400,units = 'cm')
# p3
# dev.off()

## See thcoe catch per country
c.country <- read.csv('data/catch_per_country.csv')
cps <- read.csv('data/catch_per_sector.csv')

## Calculate d

cps.s <- melt(cps, id.vars = c('year','nfish','nhauls','Country','Catch'),
              measure.vars = rep(paste('X',1:15, sep =''))) # Omit fleet from this df

cps.s <- cps.s %>% 
  group_by(year,Country, variable) %>% 
  summarise(agecomp =weighted.mean(value,Catch))

# Make ages numbers rather than factors
ages <- as.numeric(unlist(strsplit(as.character(cps.s$variable), split = "X")))
ages <- ages[is.na(ages) == 0]

cps.s$age <- ages

cps.am <- cps.s %>% 
  group_by(year,Country) %>% 
  summarise(am = sum((agecomp/100)*age))

cps.am.all <- matrix(NA, df$nyear)

for(i in 1:df$nyear){
  cps.am.all[i] <- sum(df$age[2:16]*df$age_catch[,i])
}

df.am.all <- data.frame(year = df$years, Country = 'All',am = cps.am.all)


p1 <- ggplot(cps.am, aes(x= year, y= am, color = Country, group = Country))+geom_line()+geom_point()+theme_classic()+
  geom_line(data = df.am.catch, linetype = 2)+
  scale_x_continuous(limit = c(1990,2017))+scale_y_continuous(limit = c(2.5,10),name = 'average age in catch')+
  geom_line(data = df.am.all, col = 'red')+geom_point(data=df.am.all, col ='red')

#png('age_comps.png', width = 16, height = 12, res = 400,units = 'cm')
p1
#dev.off()

# Is my weighted calculation correct? 
cps.all<- melt(cps, id.vars = c('year','nfish','nhauls','Country','Catch'),measure.vars = rep(paste('X',1:15, sep =''))) # Omit fleet from this df
cps.all <- cps.all %>% 
  group_by(year, variable) %>% 
  summarise(agecomp =weighted.mean(value,Catch))
# Make ages numbers rather than factors
ages <- as.numeric(unlist(strsplit(as.character(cps.all$variable), split = "X")))
ages <- ages[is.na(ages) == 0]

cps.all$age <- ages

cps.all.s <- cps.all %>% 
  group_by(year) %>% 
  summarise(am = sum((agecomp/100)*age))




```

## Including Plots

